Contextuality and the Born rule in fuzzy quantum theories

### 2017-04-10 prop_letter.tex
prop_letter.tex \paragraph{The Kochen-Specker Theorem} Updating...

\documentclass[english,reprint, aps, prl,superscriptaddress, showpacs,
showkeys, longbibliography, amsmath, amssymb]{revtex4-1} 
\usepackage{fullpage}
\usepackage[T1,T2A]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{babel}
\usepackage{amsthm}
\usepackage{mathrsfs}
\usepackage{bbold}
\usepackage{wesa}
\usepackage{graphicx}
\usepackage{verbatim}
\usepackage{hyperref}

@@

It asserts that there exists no function~$v$~\footnote{\yutsung{Kochen-Specker's original paper~\cite{kochenspecker1967}
used $\mathcal{E}$, but the Stanford Encyclopedia of Philosophy~\cite{Held2016}
used $v$. I think it is less confusion to use $v$ here.}} assigning definite values to all observables such that for any sequence
of commuting observables, $\left\{ \mathbf{O}_{j}\right\} $,\begin{subequations}\label{eq:Kochen-Specker-Rules}
\begin{eqnarray}
 &  & \sum_{j}v\left(\mathbf{O}_{j}\right)=v\left(\sum_{j}\mathbf{O}_{j}\right)\textrm{ and}\\
 &  & \prod_{j}v\left(\mathbf{O}_{j}\right)=v\left(\prod_{j}\mathbf{O}_{j}\right).
\end{eqnarray}
\end{subequations}Suppose that there is a function~$v$ satisfying
these two properties, it must send an observable to its eigenvalues
so that the following observables~\cite{Mermin1990Simple,peres1995quantum}:
\begin{equation}
\begin{aligned}\mathbb{1} & \otimes\sigma_{z} & \sigma_{z} & \otimes\mathbb{1} & \sigma_{z} & \otimes\sigma_{z}\\
\sigma_{x} & \otimes\mathbb{1} & \mathbb{1} & \otimes\sigma_{x} & \sigma_{x} & \otimes\sigma_{x}\\
\sigma_{x} & \otimes\sigma_{z} & \sigma_{z} & \otimes\sigma_{x} & \sigma_{y} & \otimes\sigma_{y}
\end{aligned}
\label{eq:MerminSquare}
\end{equation}
must be sent to $1$ or $-1$. In each row and column, operators commutes,
and each operator is the product of the two others, except in the
third column $\left(\sigma_{z}\otimes\sigma_{z}\right)\left(\sigma_{x}\otimes\sigma_{x}\right)=-\sigma_{y}\otimes\sigma_{y}$.
Therefore, it is clear impossible to find a function~$v$ such that
\begin{equation}
\begin{aligned}v(\mathbb{1} & \otimes\sigma_{z}) & v(\sigma_{z} & \otimes\mathbb{1}) & v(\sigma_{z} & \otimes\sigma_{z})\\
v(\sigma_{x} & \otimes\mathbb{1}) & v(\mathbb{1} & \otimes\sigma_{x}) & v(\sigma_{x} & \otimes\sigma_{x})\\
v(\sigma_{x} & \otimes\sigma_{z}) & v(\sigma_{z} & \otimes\sigma_{x}) & v(\sigma_{y} & \otimes\sigma_{y})
\end{aligned}
\label{eq:MerminSquare-values}
\end{equation}
satisfying equation~(\ref{eq:Kochen-Specker-Rules}).

\begin{table*}
\caption{\label{table:commonBasis}In each row, three observables are listed
in the left column commutes, and their common eigenbasis is listed
in the right column. In order to unify and simplify the notation,
we do not normalize eigenstates, and represent the state~$\alpha_{00}\ket{00}+\alpha_{01}\ket{01}+\alpha_{10}\ket{10}+\alpha_{11}\ket{11}$
as $\left(\alpha_{00},\alpha_{01},\alpha_{10},\alpha_{11}\right)$.
Also, to simplify the notation later, the projector onto each eigenstate
is on the right of each eigenstate. Only the projectors used in Eq.~(\ref{eq:CEGA})
will be named, and they are named so that if two projectors share
the same index, they are commuting.}

\begin{ruledtabular}
\begin{tabular}{ccc|cccc}
\multicolumn{3}{c}{Commuting observables} & \multicolumn{4}{c}{Common eigenbasis}\tabularnewline
\hline 
$\mathbb{1}\otimes\sigma_{z}$  & $\sigma_{z}\otimes\mathbb{1}$  & $\sigma_{z}\otimes\sigma_{z}$  & $(1,0,0,0)\rightarrow P_{12}$  & $(0,1,0,0)\rightarrow P_{18}$  & $(0,0,1,0)$  & $(0,0,0,1)\rightarrow P_{28}$ \tabularnewline
$\sigma_{x}\otimes\mathbb{1}$  & $\mathbb{1}\otimes\sigma_{x}$  & $\sigma_{x}\otimes\sigma_{x}$  & $(1,1,1,1)\rightarrow P_{56}$  & $(1,-1,1,-1)\rightarrow P_{59}$  & $(1,1,-1,-1)\rightarrow P_{69}$  & $(1,-1,-1,1)$ \tabularnewline
$\sigma_{x}\otimes\sigma_{z}$  & $\sigma_{z}\otimes\sigma_{x}$  & $\sigma_{y}\otimes\sigma_{y}$  & $(1,1,1,-1)\rightarrow P_{37}$  & $(1,1,-1,1)\rightarrow P_{47}$  & $(1,-1,1,1)$  & $(-1,1,1,1)\rightarrow P_{34}$ \tabularnewline
$\mathbb{1}\otimes\sigma_{z}$  & $\sigma_{x}\otimes\mathbb{1}$  & $\sigma_{x}\otimes\sigma_{z}$  & $(1,0,1,0)\rightarrow P_{48}$  & $(0,1,0,1)$  & $(1,0,-1,0)\rightarrow P_{58}$  & $(0,1,0,-1)\rightarrow P_{45}$ \tabularnewline
$\sigma_{z}\otimes\mathbb{1}$  & $\mathbb{1}\otimes\sigma_{x}$  & $\sigma_{z}\otimes\sigma_{x}$  & $(1,1,0,0)$  & $(1,-1,0,0)\rightarrow P_{67}$  & $(0,0,1,1)\rightarrow P_{17}$  & $(0,0,1,-1)\rightarrow P_{16}$ \tabularnewline
$\sigma_{z}\otimes\sigma_{z}$  & $\sigma_{x}\otimes\sigma_{x}$  & $\sigma_{y}\otimes\sigma_{y}$  & $(0,1,1,0)\rightarrow P_{29}$  & $(0,1,-1,0)\rightarrow P_{23}$  & $(1,0,0,1)\rightarrow P_{39}$  & $(1,0,0,-1)$ \tabularnewline
\end{tabular}
\end{ruledtabular}

@@

Since observables in equation~(\ref{eq:MerminSquare}) are not projections,
if we want to reformulate the above argument in the language of quantum
IVPMs, we first need to find the common eigenbasis of each row and
column, and consider the projections to each eigenvectors~\cite{peres1995quantum,CabelloEstebaranzGarcia-Alcaine1996,Cabello_2008}.
Let $\mathcal{T}_{18}$ denote the named projectors onto eigenstates
listed in Table~\ref{table:commonBasis}, and the eigenstates they
project to are denoted by $\events_{18}$. The Kochen-Specker theorem
could then be proved by showing there is no function~$v:\events_{18}\rightarrow\left\{ 0,1\right\} $
satisfying equation~(\ref{eq:Kochen-Specker-Rules}), where
\begin{equation}
\events_{18}=\set{\sum_{i=1}^{N}P_{i}}{\textrm{orthogonal }\left\{ P_{i}\right\} _{i=1}^{N}\subseteq\mathcal{T}_{24}}
\end{equation}
is a sub-event space.

Assume there is a function~$v:\events_{18}\rightarrow\left\{ 0,1\right\} $
satisfying equation~(\ref{eq:Kochen-Specker-Rules}). There must
be a quantum IVPM~$\bar{\mu}:\events_{18}\rightarrow\mathscr{I}_{\mathrm{D}}$
defined by $\bar{\mu}\left(P\right)=\iota_{\mathrm{D}}\left(v\left(P\right)\right)$,
where 
\begin{equation}
\iota_{\mathrm{D}}\left(x\right)=\begin{cases}
\necess & \text{, if }x=1\text{ ;}\\
\imposs & \text{, if }x=0\text{ .}
\end{cases}\label{eq:homomorphism-0}
\end{equation}
Because the right-hand side of each equation forms an orthogonal basis,
we can apply Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
and have the following conditions 
\begin{equation}
\begin{aligned}\necess & \subseteq\bar{\mu}\left(P_{12}\right)+\bar{\mu}\left(P_{18}\right)+\bar{\mu}\left(P_{17}\right)+\bar{\mu}\left(P_{16}\right)\textrm{ ,}\\
\necess & \subseteq\bar{\mu}\left(P_{16}\right)+\bar{\mu}\left(P_{67}\right)+\bar{\mu}\left(P_{69}\right)+\bar{\mu}\left(P_{56}\right)\textrm{ ,}\\
\necess & \subseteq\bar{\mu}\left(P_{56}\right)+\bar{\mu}\left(P_{59}\right)+\bar{\mu}\left(P_{58}\right)+\bar{\mu}\left(P_{45}\right)\textrm{ ,}\\
\necess & \subseteq\bar{\mu}\left(P_{45}\right)+\bar{\mu}\left(P_{48}\right)+\bar{\mu}\left(P_{47}\right)+\bar{\mu}\left(P_{34}\right)\textrm{ ,}\\
\necess & \subseteq\bar{\mu}\left(P_{34}\right)+\bar{\mu}\left(P_{37}\right)+\bar{\mu}\left(P_{39}\right)+\bar{\mu}\left(P_{23}\right)\textrm{ ,}\\
\necess & \subseteq\bar{\mu}\left(P_{23}\right)+\bar{\mu}\left(P_{29}\right)+\bar{\mu}\left(P_{28}\right)+\bar{\mu}\left(P_{12}\right)\textrm{ ,}\\
\necess & \subseteq\bar{\mu}\left(P_{28}\right)+\bar{\mu}\left(P_{18}\right)+\bar{\mu}\left(P_{58}\right)+\bar{\mu}\left(P_{48}\right)\textrm{ ,}\\
\necess & \subseteq\bar{\mu}\left(P_{47}\right)+\bar{\mu}\left(P_{37}\right)+\bar{\mu}\left(P_{17}\right)+\bar{\mu}\left(P_{67}\right)\textrm{ ,}\\
\necess & \subseteq\bar{\mu}\left(P_{69}\right)+\bar{\mu}\left(P_{59}\right)+\bar{\mu}\left(P_{39}\right)+\bar{\mu}\left(P_{29}\right)\textrm{ ,}
\end{aligned}
\label{eq:CEGA}
\end{equation}
If we sum up the preceding equations, we will have
\begin{equation}
\left[9,9\right]\subseteq2\sum_{\ket{\psi}\in\mathcal{T}_{18}}\bar{\mu}\left(\ket{\psi}\right)\textrm{ .}\label{eq:QuantumInterval-valuedProbability-Total}
\end{equation}
However, Eq.~(\ref{eq:QuantumInterval-valuedProbability-Total})
can never be satisfied. Because each term in right-hand side is either
$\imposs=\left[0,0\right]$ or $\necess=\left[1,1\right]$, the right-hand
side is a singleton set of an even integer while the left-hand side
is $\left[9,9\right]$. Therefore, there is no quantum IVPM~$\bar{\mu}:\events_{18}\rightarrow\mathscr{I}_{\mathrm{D}}$.
Moreover, there is no quantum IVPM defined on the set of all projectors
as well; otherwise, if we have a quantum IVPM~$\bar{\mu}':\events\rightarrow\mathscr{I}_{\mathrm{D}}$,
then we can restrict $\bar{\mu}'$ and define $\bar{\mu}=\left.\bar{\mu}'\right|_{\events_{18}}$.

>>>>

\documentclass[english,reprint, aps, prl,superscriptaddress, showpacs,
showkeys, longbibliography, amsmath, amssymb]{revtex4-1} 
\usepackage{fullpage}
\usepackage[T1,T2A]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{babel}
\usepackage{mmap}
\usepackage{amsthm}
\usepackage{mathrsfs}
\usepackage{bbold}
\usepackage{wesa}
\usepackage{graphicx}
\usepackage{verbatim}
\usepackage{hyperref}

@@

It asserts that there exists no function~$v$~\footnote{\yutsung{Kochen-Specker's original paper~\cite{kochenspecker1967}
used $\mathcal{E}$, but the Stanford Encyclopedia of Philosophy~\cite{Held2016}
used $v$. I think it is less confusion to use $v$ here.}} assigning definite values to all observables such that for any sequence
of commuting observables, $\left\{ \mathbf{O}_{j}\right\} $,\begin{subequations}\label{eq:Kochen-Specker-Rules}
\begin{eqnarray}
 &  & \sum_{j}v\left(\mathbf{O}_{j}\right)=v\left(\sum_{j}\mathbf{O}_{j}\right)\textrm{ and}\\
 &  & \prod_{j}v\left(\mathbf{O}_{j}\right)=v\left(\prod_{j}\mathbf{O}_{j}\right).
\end{eqnarray}
\end{subequations}Suppose that there is a function~$v$ satisfying
these two properties, it must send an observable to its eigenvalues
so that the following observables~\cite{Mermin1990Simple,peres1995quantum}:
\begin{equation}
\begin{aligned}\mathbb{1} & \otimes\sigma_{z} & \sigma_{z} & \otimes\mathbb{1} & \sigma_{z} & \otimes\sigma_{z}\\
\sigma_{x} & \otimes\mathbb{1} & \mathbb{1} & \otimes\sigma_{x} & \sigma_{x} & \otimes\sigma_{x}\\
\sigma_{x} & \otimes\sigma_{z} & \sigma_{z} & \otimes\sigma_{x} & \sigma_{y} & \otimes\sigma_{y}
\end{aligned}
\label{eq:MerminSquare}
\end{equation}
must be sent to $1$ or $-1$. In each row and column, operators commutes,
and each operator is the product of the two others, except in the
third column $\left(\sigma_{z}\otimes\sigma_{z}\right)\left(\sigma_{x}\otimes\sigma_{x}\right)=-\sigma_{y}\otimes\sigma_{y}$.
Therefore, it is clear impossible to find a function~$v$ such that
\begin{equation}
\begin{aligned}v(\mathbb{1} & \otimes\sigma_{z}) & v(\sigma_{z} & \otimes\mathbb{1}) & v(\sigma_{z} & \otimes\sigma_{z})\\
v(\sigma_{x} & \otimes\mathbb{1}) & v(\mathbb{1} & \otimes\sigma_{x}) & v(\sigma_{x} & \otimes\sigma_{x})\\
v(\sigma_{x} & \otimes\sigma_{z}) & v(\sigma_{z} & \otimes\sigma_{x}) & v(\sigma_{y} & \otimes\sigma_{y})
\end{aligned}
\label{eq:MerminSquare-values}
\end{equation}
satisfying Eq.~(\ref{eq:Kochen-Specker-Rules}).

\begin{table*}
\caption{\label{table:commonBasis}In each row, three observables are listed
in the left column commutes, and their common eigenbasis is listed
in the right column. In order to unify and simplify the notation,
we do not normalize eigenstates, and represent the state~$\alpha_{00}\ket{00}+\alpha_{01}\ket{01}+\alpha_{10}\ket{10}+\alpha_{11}\ket{11}$
as $\left(\alpha_{00},\alpha_{01},\alpha_{10},\alpha_{11}\right)$.
Also, to simplify the notation further, the projector onto each eigenstate
used in Eq.~(\ref{eq:CEGA}) is named such that if two projectors
share the same index, they are commuting.}

\begin{ruledtabular}
\begin{tabular}{c|c}
Commuting observables & Common eigenbasis\tabularnewline
\hline 
$\begin{aligned}\mathbb{1} & \otimes\sigma_{z} & \sigma_{z} & \otimes\mathbb{1} & \sigma_{z} & \otimes\sigma_{z}\\
\sigma_{x} & \otimes\mathbb{1} & \mathbb{1} & \otimes\sigma_{x} & \sigma_{x} & \otimes\sigma_{x}\\
\sigma_{x} & \otimes\sigma_{z} & \sigma_{z} & \otimes\sigma_{x} & \sigma_{y} & \otimes\sigma_{y}\\
\mathbb{1} & \otimes\sigma_{z} & \sigma_{x} & \otimes\mathbb{1} & \sigma_{x} & \otimes\sigma_{z}\\
\sigma_{z} & \otimes\mathbb{1} & \mathbb{1} & \otimes\sigma_{x} & \sigma_{z} & \otimes\sigma_{x}\\
\sigma_{z} & \otimes\sigma_{z} & \sigma_{x} & \otimes\sigma_{x} & \sigma_{y} & \otimes\sigma_{y}
\end{aligned}
$ & $\begin{aligned}(1,0,0,0) & \rightarrow P_{12} & (0,1,0,0) & \rightarrow P_{18} & (0,0,1,0) &  & (0,0,0,1) & \rightarrow P_{28}\\
(1,1,1,1) & \rightarrow P_{56} & (1,-1,1,-1) & \rightarrow P_{59} & (1,1,-1,-1) & \rightarrow P_{69} & (1,-1,-1,1)\\
(1,1,1,-1) & \rightarrow P_{37} & (1,1,-1,1) & \rightarrow P_{47} & (1,-1,1,1) &  & (-1,1,1,1) & \rightarrow P_{34}\\
(1,0,1,0) & \rightarrow P_{48} & (0,1,0,1) &  & (1,0,-1,0) & \rightarrow P_{58} & (0,1,0,-1) & \rightarrow P_{45}\\
(1,1,0,0) &  & (1,-1,0,0) & \rightarrow P_{67} & (0,0,1,1) & \rightarrow P_{17} & (0,0,1,-1) & \rightarrow P_{16}\\
(0,1,1,0) & \rightarrow P_{29} & (0,1,-1,0) & \rightarrow P_{23} & (1,0,0,1) & \rightarrow P_{39} & (1,0,0,-1)
\end{aligned}
$\tabularnewline
\end{tabular}
\end{ruledtabular}

@@

Since observables in Eq.~(\ref{eq:MerminSquare}) are not projections,
if we want to reformulate the above argument in the language of IVPMs,
we first need to find the common eigenbasis of each row and column,
and consider the projections to each eigenvectors~\cite{peres1995quantum,CabelloEstebaranzGarcia-Alcaine1996,Cabello_2008}.
Let $\mathcal{T}_{18}$ denote the named projectors onto eigenstates
listed in Table~\ref{table:commonBasis}, and the eigenstates they
project to are denoted by $\events_{18}$. The Kochen-Specker theorem
could then be proved by showing there is no function~$v:\events_{18}\rightarrow\left\{ 0,1\right\} $
satisfying Eq.~(\ref{eq:Kochen-Specker-Rules}), where
\begin{equation}
\events_{18}=\set{\sum_{i=1}^{N}P_{i}}{\textrm{orthogonal }\left\{ P_{i}\right\} _{i=1}^{N}\subseteq\mathcal{T}_{18}}
\end{equation}
is a sub-event space.

Assume there is a function~$v:\events_{18}\rightarrow\left\{ 0,1\right\} $
satisfying Eq.~(\ref{eq:Kochen-Specker-Rules}). There must be a
quantum IVPM~$\bar{\mu}:\events_{18}\rightarrow\mathscr{I}_{\mathrm{D}}$
defined by $\bar{\mu}\left(P\right)=\iota_{\mathrm{D}}\left(v\left(P\right)\right)$,
where 
\begin{equation}
\iota_{\mathrm{D}}\left(x\right)=\begin{cases}
\necess & \text{, if }x=1\text{ ;}\\
\imposs & \text{, if }x=0\text{ .}
\end{cases}\label{eq:homomorphism-0}
\end{equation}
Since each orthogonal basis can form an instance of Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion}),
the orthogonal bases in $\events_{18}$ gives the following equations:
\begin{equation}
\begin{aligned}\necess & \subseteq\bar{\mu}\left(P_{12}\right)+\bar{\mu}\left(P_{18}\right)+\bar{\mu}\left(P_{17}\right)+\bar{\mu}\left(P_{16}\right)\textrm{ ,}\\
\necess & \subseteq\bar{\mu}\left(P_{16}\right)+\bar{\mu}\left(P_{67}\right)+\bar{\mu}\left(P_{69}\right)+\bar{\mu}\left(P_{56}\right)\textrm{ ,}\\
\necess & \subseteq\bar{\mu}\left(P_{56}\right)+\bar{\mu}\left(P_{59}\right)+\bar{\mu}\left(P_{58}\right)+\bar{\mu}\left(P_{45}\right)\textrm{ ,}\\
\necess & \subseteq\bar{\mu}\left(P_{45}\right)+\bar{\mu}\left(P_{48}\right)+\bar{\mu}\left(P_{47}\right)+\bar{\mu}\left(P_{34}\right)\textrm{ ,}\\
\necess & \subseteq\bar{\mu}\left(P_{34}\right)+\bar{\mu}\left(P_{37}\right)+\bar{\mu}\left(P_{39}\right)+\bar{\mu}\left(P_{23}\right)\textrm{ ,}\\
\necess & \subseteq\bar{\mu}\left(P_{23}\right)+\bar{\mu}\left(P_{29}\right)+\bar{\mu}\left(P_{28}\right)+\bar{\mu}\left(P_{12}\right)\textrm{ ,}\\
\necess & \subseteq\bar{\mu}\left(P_{28}\right)+\bar{\mu}\left(P_{18}\right)+\bar{\mu}\left(P_{58}\right)+\bar{\mu}\left(P_{48}\right)\textrm{ ,}\\
\necess & \subseteq\bar{\mu}\left(P_{47}\right)+\bar{\mu}\left(P_{37}\right)+\bar{\mu}\left(P_{17}\right)+\bar{\mu}\left(P_{67}\right)\textrm{ ,}\\
\necess & \subseteq\bar{\mu}\left(P_{69}\right)+\bar{\mu}\left(P_{59}\right)+\bar{\mu}\left(P_{39}\right)+\bar{\mu}\left(P_{29}\right)\textrm{ .}
\end{aligned}
\label{eq:CEGA}
\end{equation}
If we sum up the preceding equations, we will have
\begin{equation}
\left[9,9\right]\subseteq2\sum_{\ket{\psi}\in\mathcal{T}_{18}}\bar{\mu}\left(\ket{\psi}\right)\textrm{ .}\label{eq:QuantumInterval-valuedProbability-Total}
\end{equation}
However, Eq.~(\ref{eq:QuantumInterval-valuedProbability-Total})
can never be satisfied. Because each term in right-hand side is either
$\imposs=\left[0,0\right]$ or $\necess=\left[1,1\right]$, the right-hand
side is a singleton set of an even integer while the left-hand side
is $\left[9,9\right]$. Therefore, there is no quantum IVPM~$\bar{\mu}:\events_{18}\rightarrow\mathscr{I}_{\mathrm{D}}$.
Moreover, there is no quantum IVPM defined on the set of all projectors
as well; otherwise, if we have a quantum IVPM~$\bar{\mu}':\events\rightarrow\mathscr{I}_{\mathrm{D}}$,
then we can restrict $\bar{\mu}'$ and define $\bar{\mu}=\left.\bar{\mu}'\right|_{\events_{18}}$.
